---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ServicePlanningFunctions

<!-- badges: start -->

<!-- badges: end -->

The goal of ServicePlanningFunctions is to standardize functions repeatedly used in service planning analyses at King County Metro. These are the functions that have been developed over multiple analyses and projects, so they are not all necessarily related to each other. Read through the function descriptions below to learn about available functions

## Installation

You can install the development version of ServicePlanningFunctions from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("KC-Metro-Transit/ServicePlanningFunctions")
```

## save_objects() Example

save_objects() is designed to work with the purrr package to iterate over all objects in your environment and save a copy of them to disk. This is useful if you are trying to save your work or if you are getting data ready to import into a Shiny app.

```{r example_save_objects, eval=F}
library(ServicePlanningFunctions)
## basic example code
object_1 <- cars
object_2 <- iris
object_3 <- gapminder

object_list <- ls() #list all object names you want to export


object_list <- object_list[! object_list %in% c("object_list")] #remove the list of objects from the list of names of objects to export

purrr::map(object_list, file_location ="your/file/location",  save_objects) #choose where the files should write to disk


```

## remove_water() Example

```{r example_water, eval=F}
 kc_tracts <- sf::read_sf(fs::path_package( "extdata", "2020_Census_Tracts_for_King_County___tracts20_area.shp", package = "ServicePlanningFunctions"))

kc_tracts_no_water <- remove_water(polygon = kc_tracts, state_code = "WA", county_code = "King", crs = 2926)

mapview::mapview(kc_tracts_no_water)
```

## combine_gtfs() Example
combine_gtfs is designed to be used with GTFS files from Netplan and combines multiple sets of GTFS files into one. This is typically used to combine weekday, saturday and sunday sets of GTFS files into one.
```{r example_combine_gtfs, eval=F}

gtfs_folder <- fs::path_package( "extdata", "combine_gtfs", package = "ServicePlanningFunctions")

madison_st_area_gtfs <- combine_gtfs(gtfs_filepath = gtfs_folder, designated_start_date = 20240901, designated_end_date = 20250330, save_csv = FALSE, save_RDS = FALSE, output_folder = NULL)


```

## append_gtfs() Example
append_gtfs is designed to be used with two sets of GTFS files from Netplan: (1) One set of GTFS files from the ***Proposed Network that only contains select routes*** and (2) one set of GTFS files from the ***Baseline Network containing all routes***. The function will produce a new GTFS output that represents the full proposed network that contains proposed changes for the selected routes and the baseline network for the remaining routes.

This can later be used as a valid input when performing trip change analysis, accessibility analysis, and the Title VI analysis. Users also need to explicitly call out deleted routes, so that they are not part of the final output. It is also recommended that the *combine_gtfs* is run beforehand for each network.

## combine_interagency_gtfs
combine_interagency_gtfs is designed to combine the NetPlan GTFS of Metro's full network and GTFS files from other agencies: Sound Transit, Community Transit, and Pierce Transit. Service date ranges of other agencies should overlap the service date range of Metro's GTFS. It is also recommended that the *combine_gtfs* is run beforehand for Metro's GTFS. If Metro's GTFS already contain routes from other agencies, they are dropped and replaced with that from their respective agency.

## clean_service_rte_num() Example
clean_service_rte_num adds the service_rte_num field to the routes.txt GTFS file that represents the number equivalent of the route name as indicated by the route_short_name field. If the function does not recognize a route name, it will prompt the user to provide a number to represent the route name.
```{r example_clean_service_rte_num, eval=F}


spring_24_gtfs <- tidytransit::read_gtfs(path =  fs::path_package( "extdata", "gtfs", "241_gtfs.zip", package = "ServicePlanningFunctions"))

spring_24_routes <- clean_service_rte_num(spring_24_gtfs$routes, netplan_gtfs = FALSE)

netplan_sept24_gtfs <- tidytransit::read_gtfs(path =  fs::path_package( "extdata", "gtfs", "SEPT24_TRIP_GTFS.zip", package = "ServicePlanningFunctions"))

netplan_routes <- clean_service_rte_num(netplan_sept24_gtfss$routes, netplan_gtfs = TRUE)


```

## gtfs_calendar_full_dates() Example

```{r example_gtfs_calendar_full_dates, eval=F}

spring_24_gtfs <- tidytransit::read_gtfs(path =  fs::path_package( "extdata", "gtfs", "241_gtfs.zip", package = "ServicePlanningFunctions"))

full_cal <- gtfs_calendar_full_dates(calendar = spring_24_gtfs$calendar, calendar_dates = spring_24_gtfs$calendar_dates, netplan_gtfs = FALSE)

netplan_sept24_gtfs <- tidytransit::read_gtfs(path =  fs::path_package( "extdata", "gtfs", "SEPT24_TRIP_GTFS.zip", package = "ServicePlanningFunctions"))

netplan_full_cal <- gtfs_calendar_full_dates(
                            calendar = netplan_sept24_gtfs$calendar,                                           calendar_dates = netplan_sept24_gtfs$calendar_dates, 
                            netplan_gtfs = TRUE, 
                            designated_start_date = 20240901,                                                  designated_end_date = 20250330)

```

## st_erase() Example

A helper function for remove_water() that clips a polygon out from another polygon. Results in a "donut" polygon.

```{r example_st_erase, eval=F}
 big_box <- sf::st_polygon(x= list(polygon))
 plot(sf::st_geometry(big_box))
 shape_to_remove <- matrix(c(1,1,1,2,2,2,2,1,1,1),ncol=2, byrow=TRUE)
#create smaller polygon to subset
 polygon_to_remove <- sf::st_polygon(x = list(shape_to_remove))
 plot(sf::st_geometry(polygon_to_remove))
# Polygon with smaller area removed
 clipped_polygon <- st_erase(big_box, polygon_to_remove)
 plot(st_geometry(clipped_polygon))
```

## count_trips_by_geography() Example

Function takes a GTFS and a spatial dataset and counts unique trips at stops in each polygon. 
